Deciding to Disclose: Pregnancy and Alcohol Misuse

### 2014-07-16 Background.tex
Neural correlates of decisions

@@

\citet{Gigerenzer1996} take the concept of bounded rationality further, and argue for what they term \ac{FFH}. This recasts rationality as bound to the context of the behaviour - a rational approach to choosing the right mate might well require checking every possible partner, but given finite time, memory, and so on rapidly becomes nonsensical.

@@

While heuristic theories arguably fall under the purview of the descriptive, the wider tendency is towards what are in essence patches to normative models. The most influential models in this class derive from \ac{PT} \citep{Kahneman1979}, which combines a set of heuristics based on observed decision behaviour, with distortions to the perception of probability, and the value of outcomes. \citet{Tversky1992} subsequently addressed key issues present in their original formulation by introducing \ac{CPT}, which allows for non-binary decisions, at the expense of the heuristics. The essence then, is that high and low probabilities are treated differently, and the subjective value of a loss differs from the equivalent gain (losing your shirt is worse than winning a shirt is good).

@@

While vastly influential, the expected utility theory has been substantially
criticised, generally for failing to predict real behaviour. \citet{Society2013}
attacked the independence axiom in particular, suggesting that in
some scenarios people's choices would be inconsistent where expected
utility implies otherwise. A number of studies (e.g. \citep{Oliver2003,Burke1996})
have since supported the intuition to some extent.

More recently, support for some aspects of the expected utility theory,
particularly the concept of utility as a common currency for comparison,
has come from neurology, for example following work by \citet{Platt1999},
\citet{Padoa-Schioppa2006,Padoa-Schioppa2008} report neuronal firing
corresponding to economic value in decision making tasks undertaken
by monkeys, while \citet{Christopoulos2009} found similarly indicative
results for risk aversion. The suggestion implicit in the model proposed
here, that this also applies to social judgements, is less investigated,
although both \citet{Watson2012}, and \citet{Willis2010} found that
lesions in the brain area%
\footnote{The orbitofrontal cortex.%
} identified by \citeauthor{Padoa-Schioppa2006} lead to abnormal social
judgements in humans and primates.

>>>>

Recently, several studies have explored biological correlates to aspects of expected utility. The fundamental concept, that all outcomes are comparable in a universal currency has been supported by evidence of neural correlates of decision variables \citep{Platt1999}, and following from this results from \citet{Padoa-Schioppa2006,Padoa-Schioppa2008} showing neuronal firing in the \ac{OFC} corresponding to revealed preferences in monkeys. Additionally, some support for neural representation of value, and risk aversion was found by \citet{Christopoulos2009}. The model presented in this paper makes an explicit assumption that social decisions utilise the same process, and while this is less well supported there is some evidence to suggest involvement by the same brain region, since damage to the \ac{OFC} has been shown to impair social judgements in both primates \citep{Watson2012}, and humans \citep{Willis2010}.

@@

\citet{Gigerenzer1996} take the concept of bounded rationality further, and argue for what they term \ac{FFH}. This recasts rationality as bound to the context of the behaviour - a rational approach to choosing the right mate might well require checking every possible partner, but given finite time, memory, and so on rapidly becomes nonsensical. On this basis, they contend that the rationality of any given decision rule can only be determined in the context of the environment, which implies that heuristics are task specific.

@@

While heuristic theories arguably fall under the purview of the descriptive, the wider tendency is towards what are in essence patches to normative models. The most influential models in this class derive from \ac{PT} \citep{Kahneman1979}, which combines a set of heuristics based on observed decision behaviour, with distortions to the perception of probability, and the value of outcomes. \citet{Tversky1992} subsequently addressed key issues present in their original formulation by introducing \ac{CPT}, which allows for non-binary decisions, at the expense of the heuristics. The essence then, is that high and low probabilities are treated differently, and the subjective value of a loss differs from the equivalent gain (losing your shirt is perceived as more of a loss than winning a shirt is a gain).

@@
